Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar

### 2021-02-01 FS-CRB-HR-report1.tex
final

\begin{footnotesize}
	\todo{change url}
\url{https://github.com/aubreymoore/FY19-PPA-Report-1/raw/master/PPA19-report2.pdf}
\end{footnotesize}

@@

In a previous Forest Service feasability study, we successfully tracked CRB to cryptic breeding sites using miniature radio transmitters. However, the high cost of radio tags (>\$100 each) and short shelf life and field life of nonreplacable batteries (about 1 month and one week, respectively) preclude radio tracking for routine CRB breeding site detection make this technique too expensive for routine surveys.

A promising alternative is harmonic radar technology. Harmonic radar tags do not require a battery. They are inexpensive (about \$1 each) and they have unlimited shelf-life and field life. We are evaluating hand-held harmonic radar equipment and tags manufactured by RECCO in Sweden. Rescuers use the hand-held units to rapidly locate victims wearing tags sown into their clothing. RECCO technology has been used to track other insects but has not yet been tried with CRB.

Our objective is not to track CRB tagged beetles in real-time, but to discover the end points of tags several days or even weeks after release of the tagged beetles.We anticipate that tags will accumulate at breeding sites.

Here is the plan from the \textit{verbatim} from the approved grant proposal:

@@

The most important activity is \textbf{Conduct field releases and tracking of tagged CRB (two week intensive fieldwork period)}. The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar (similar to what we did in the radio-tracking feasibility study). This activity has not yet been scheduled because of COVID-19 travel restrictions \ref{impediments}. Otherwise we are prepared to proceed.

@@

Progress on this project was impeded by COVID-19 travel restrictions which prevented collaborators from visiting Guam to participate in field work. Further delayed by delay was caused by Government of Guam \textit{stay at home} orders. The University of Guam was officially closed from March 20 to May 10 2020 and again from August 16 2020 to January 15 2021.

>>>>

\begin{footnotesize}
\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report1.pdf}
\end{footnotesize}

@@

Previously, we successfully tracked CRB to cryptic breeding sites using miniature radio transmitters. However, the high cost of radio tags (>\$100 each) and short shelf life and field life of nonreplacable batteries (about 1 month and one week, respectively) make this technique too expensive for routine surveys to detect CRB breeding sites.

A promising alternative technology is harmonic radar. Harmonic radar tags do not require a battery. They are inexpensive (about \$1 each) and they have unlimited shelf-life and field life. We are evaluating hand-held harmonic radar equipment and tags manufactured by RECCO in Sweden. Rescuers use the hand-held units to rapidly locate victims wearing tags sown into their clothing. RECCO technology has been used to track other insects but has not yet been tried with CRB.

Our objective is not to track CRB tagged beetles in real-time, but to discover the end points of tags several days or even weeks after release of tagged beetles. We anticipate that tags will accumulate at breeding sites.

Here is the plan from the approved grant proposal:

@@

The most important activity is \textbf{Conduct field releases and tracking of tagged CRB (two week intensive fieldwork period)}. The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar. (We are following an approach similar to what we did in the radio-tracking feasibility study). This fieldwork activity has not yet been scheduled because of COVID-19 travel restrictions \ref{impediments}. Otherwise we are prepared to proceed.

@@

Progress on this project was impeded by COVID-19 travel restrictions which prevented collaborators from visiting Guam to participate in field work. Further delayed by delay was caused by Government of Guam \textit{stay at home} orders.

The University of Guam was officially closed from March 20 to May 10 2020 and again from August 16 2020 to January 15 2021.
